Silent pessimists - researchers' perceptions of malaria eradication and implications for health spending

### 2017-10-05 paper.Rmd
update

```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = F}
# No scientific notation
options(scipen=999)

@@

All analysis was carried out in R language [@R]. All analysis code, as well as the code used for the identification and contacting of participants, is publicly available at [https://github.com/joebrew/malaria_survey](https://github.com/joebrew/malaria_survey).

The searching and retrieval of information from the PubMed database was carried out using the RISmed package [@Kovalchik2015]. Qualitative analysis relied on the RQDA package (citation needed here).

@@

De-biasing of sample selection was carried out using Heckman's 2 step approach (citation needed), as implemented by Toomet [@toomet2008].

@@

A total of `r nrow(df)` researchers participated in the survey.

## Sample selection and de-biasing of results

This section is unfinished. Below are regression outputs for both OLS and Heckman's 2 step model. I will update this soon.

```{r, results = 'asis'}
stargazer(ols1, heck1, ml1,
          header = FALSE,
          # se=list(cse(ols1),NULL,NULL),
          title="Regression on years until eradication", 
          type="latex",
          df=FALSE, digits=4)
# summary(ols1)
# summary(heck1)

@@

## Expertise profile

```{r}
x <- df %>%
  group_by(n_areas) %>%
  tally %>%
  mutate(p = n / sum(n) * 100)
```

@@

Participants had a total of `r length(old_areas)` unique (self-described) areas of expertise, of which `r length(areas)` areas of expertise had 5 or more participants.

```{r}
temp <- areas_df %>%
  arrange(desc(count))
temp$area <- trimws(temp$area, which = 'both')
names(temp) <- c('Area', 'Researchers')
temp$Area <- as.character(temp$Area)
temp$Area <- Hmisc::capitalize(temp$Area)
kable(temp)
```

@@

>>>>

```{r, echo = FALSE, warning = FALSE, message = FALSE, comment = NA, error= FALSE, cache = F}
# No scientific notation
# options(scipen=999)

@@

All analysis was carried out in R language [@R]. All analysis code, as well as the code used for the identification and contacting of participants, is publicly available at [https://github.com/joebrew/malaria_survey](https://github.com/joebrew/malaria_survey). The searching and retrieval of information from the PubMed database was carried out using the RISmed package [@Kovalchik2015].

@@

De-biasing of sample selection was carried out using Heckman's 2 step approach as implemented by Toomet [@toomet2008].

@@

A total of `r length(which(df$responded))` researchers participated in the survey from the `r nrow(df)` invitations sent (participation rate of `r round(length(which(df$responded)) / nrow(df) * 100, 2)`%).

## Sample selection

Respondents were qualitatively different from non-respondents. Importantly, the average number of total author-specific citations was `r round(mean(df$n_citations_total[df$responded]), digits = 2)` among respondents, but `r round(mean(df$n_citations_total[!df$responded]), digits = 2)` among non-respondents. This suggests either a tendency for more senior or more cited researchers to not respond. The distribution of author-specific total citations is shown below (a).

```{r}
g1 <- ggplot(data = df %>%
         mutate(responded = ifelse(responded, 'Respondents', 'Non-respondents')), 
       aes(x = n_citations_total)) +
  geom_density(aes(group = responded,
                   fill = responded),
               alpha = 0.6) +
  scale_x_log10() +
  theme_publication() +
  scale_fill_manual(name = '',
                    values = c('darkorange', 'lightblue')) +
  labs(x = 'Citations (log scale)',
       y = 'Density',
       title = 'Total lifetime citations')
```

When examining the number of average citations per article, the difference between respondents was slightly less: `r round(median(df$n_citations_total[df$responded]), digits = 2)` among respondents, and `r round(median(df$n_citations_total[!df$responded]), digits = 2)` among non-respondents. Again, the distribution is shown below (b).

```{r}
g2 <- ggplot(data = df %>%
         mutate(responded = ifelse(responded, 'Respondents', 'Non-respondents')), 
       aes(x = n_citations_average)) +
  geom_density(aes(group = responded,
                   fill = responded),
               alpha = 0.6,
               n = 10) +
  scale_x_log10() +
  theme_publication() +
  scale_fill_manual(name = '',
                    values = c('darkorange', 'lightblue')) +
  labs(x = 'Citations (log scale)',
       y = 'Density',
       title = 'Average citations per article')
Rmisc::multiplot(g1, g2, cols = 2)
```

The below table shows the response rate by different demographic groups of interest.

```{r}
# function to get response rate per categorical variable
response_rate <- function(variable = 'sex'){
  df$x <- df[,variable]
  out <- df %>%
    group_by(x) %>%
    summarise(responded = length(which(responded)),
              n = n()) %>%
    ungroup %>%
    mutate(p = responded / n * 100)
  return(out)
}
out_list <- list()
vars <- c('sex',
            'ethnicity',
            'bin_citations')
for (i in 1:length(vars)){
  this_var <- vars[i]
  x <- response_rate(variable = this_var)
  x$variable <- this_var
  out_list[[i]] <- x
}
out <- bind_rows(out_list) %>%
  dplyr::select(variable, 
                x,
                responded,
                n, 
                p) %>%
  rename(Variable = variable,
         Characteristic = x,
         Responded = responded,
         Invited = n,
         `Response rate` = p)
kable(out, format = 'latex', caption = 'Response rate by author characteristics')
```

The below chart shows response rate by number of citations.

```{r}
x <- out %>%
  filter(Variable == 'bin_citations') 
ggplot(data = x,
       aes(x = Characteristic,
           y = `Response rate`)) +
  geom_bar(stat = 'identity',
           fill = 'lightblue',
           alpha = 0.8) +
  theme_publication() +
  labs(x = 'Number of total citations',
       y = 'Response rate',
       title = 'Response rate by citations') +
  geom_label(aes(label = paste0(round(`Response rate`, digits = 2), '%')))
```

## De-biasing of results

```{r, results = 'asis'}
stargazer(ols1, heck1, ml1,
          header = FALSE,
          # se=list(cse(ols1),NULL,NULL),
          title="Regression on years until eradication", 
          type="text",
          df=FALSE, digits=4)
# summary(ols1)
# summary(heck1)

@@

## Expertise profile among respondents

```{r}
x <- df %>%
  filter(responded) %>%
  group_by(n_areas) %>%
  tally %>%
  mutate(p = n / sum(n, na.rm = TRUE) * 100)
```

@@

Participants had a total of `r length(old_areas)` unique (self-described) areas of expertise, of which `r length(areas)` areas of expertise had 5 or more participants (see table 1).

```{r}
temp <- areas_df %>%
  arrange(desc(count))
temp$area <- trimws(temp$area, which = 'both')
names(temp) <- c('Area', 'Researchers')
temp$Area <- as.character(temp$Area)
temp$Area <- Hmisc::capitalize(temp$Area)
kable(temp[1:10,], format = "latex", caption = 'Top 10 areas') %>%
  kableExtra::kable_styling(#position = "float_right",
                            font_size = 7)
```

@@

## Qualitative analysis of results

Participants were invited to provide "any general comments on the timeline and likelihood of global eradication". Of the `r length(which(df$responded))` who responded to the survey, `r length(which(!is.na(df$comments)))` (`r round(length(which(!is.na(df$comments))) / length(which(df$responded)) * 100, digits = 2)`%) provided a comment.
